---
title: "BOSS"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to get run BOSS model, from preprocessing to visualization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

The `bossR` package segments and tracks the cells from the BOSS method. Access to the full papers can be found [here](https://www.biorxiv.org/content/10.1101/2022.06.17.495689v1). Additionally, it allows for visualization of the segmented/tracked cells overlayed in original image.

## Installation

You can install the development version of bossR like so:

```{r, eval = FALSE}
devtools::install_github('calebae/bossR')
```

```{r, warning = FALSE, message = FALSE}
library(tiff)
library(betareg)
library(oro.nifti)
library(mmand)
library(scales)
library(dplyr)
library(parallel)
library(bossR)
```

## Tutorial Data

We will be using data from one of microarray images, published in [Remyelination alters the pattern of myelin in the cerebral cortex](https://elifesciences.org/articles/56621). Data can be obtained from [here](https://upenn.box.com/s/g5wg36fijfl5j56laoz31mn5u6s3y9kx)

The `BOSS` algorithm is designed to segment and track any cells in longitudinally obtained images. For illustration, we show how `BOSS` algorithm works to track oligodendrocytes.


### Getting the data 

`tiff` file can be loaded into `R` by using `tiff::readTIFF()` function. 

```{r}
setwd("~/Box Sync/BOSS_example/")
trt.mouse <- readTIFF("example_img.tif", all=TRUE, info = TRUE)
```

Since our data is 4D which is 3D ($X, Y, Z$) + time ($T$) whereas `nifti` file can contain 3D, subset the `nifti` files for each time-point. Total number $Z$ and $T$ can be obtained from `tiff::readTIFF()` function. 

 `tiff::readTIFF()` stores 2D ($X, Y$) matrix in every $Z \cdot T$ lists, therefore we first need to get $Z$ and $T$ from TIFF tags and transform `tiff` file to `nifti` file format with right dimension.

```{r, include=FALSE}
tiff.info <- attr(trt.mouse[[1]], "description")
Z.info <- unlist(strsplit(tiff.info, split = "\n"))[3]
Z <- as.numeric(unlist(strsplit(Z.info, split="="))[2])

T.info <- unlist(strsplit(tiff.info, split = "\n"))[4]
T <- as.numeric(unlist(strsplit(T.info, split="="))[2])

# Resoultion information will be used later
X.resolution.info <- attr(trt.mouse[[1]], "x.resolution")
Y.resolution.info <- attr(trt.mouse[[1]], "y.resolution")
```

There are $Z$ =`r Z`, $T$ = `r T`. 

Code below illustrates how to construct 3D images, from stacks of 2D `tiff` images.
Depending on the file structure, you will need to modify the code below to construct 3D images.

```{r, include=FALSE}
rotate <- function(x) t(apply(x, 2, rev))

nifti1 <- array(NA, dim=c(rev(dim(trt.mouse[[1]])[1:2]), Z))
for(i in 1:Z){
  # Note that we need to rotate 90 degree clockwise
  # to get the same image as FIJI
  nifti1[,,i] <- rotate(trt.mouse[[i]][,,1])
}

rm(trt.mouse)

trt.nifti1 <- as.nifti(nifti1)
```

Brain image of cuprizone-treated mouse at $Z$ = 100 and $T$ = 1 is as follows.
```{r} 
image(trt.nifti1, z = 21, plot.type='single')
```

### Getting the manual-segmentation for evaluation

 If you have a manual segmentation, you can overlay the segmentation on its image.
```{r}
setwd("~/Box Sync/BOSS_example/")
trt.mask<-read.csv("Stat Results Table.csv")
```

 For exploratory purpose, let's get the ROI at $T=1$.
 
 Note that original file records the ROI with its coordinate ($X, Y$), therefore we need to convert it by using its resolution information.

```{r}
trt.mask1 <- trt.mask %>% filter(T==1) 
mask.dat <- array(NA, dim = dim(trt.nifti1))
  
for(i in 1:Z){
  tmp <- trt.mask1 %>% filter(Z == i)
  if(nrow(tmp)==0){next}
  tmp.dat <- data.frame(X.pixel = tmp$X * X.resolution.info, 
                          Y.pixel = tmp$Y * Y.resolution.info,
                          Z = i)
  
  for(j in 1:nrow(tmp.dat)){
    X.pixel <- round(tmp.dat$X.pixel[j])
    X.pixel <- (X.pixel-5) : (X.pixel+5)
    Y.pixel <- round(tmp.dat$Y.pixel[j])
    Y.pixel <- (Y.pixel-5) : (Y.pixel+5)
    Z.pixel <- max(1,(tmp.dat$Z[j]-2)) : min(Z,(tmp.dat$Z[j]+2))
    # Be careful about Y axis.
    mask.dat[X.pixel, dim(trt.nifti1)[2]-Y.pixel, Z.pixel] <- 255
  }
}

mask.dat <- as.nifti(mask.dat)
```

 Image with ROI at $Z$ = 91, $T$ = 1 is as follows. This image will be used to see
 the performance of `neuroim::connComp3D`. 
 This image indicates that there could be some FALSE NEGATIVE issues when we validate our model.
 
```{r}
overlay.nifti(trt.nifti1, mask.dat, z = 21, plot.type = 'single')
```

### Intensity Modeling

We highly recommend to use the parallel computing

```{r}
roi.img <- boss_betamix(img = trt.nifti1, output_directory = paste0(getwd(),"/Binarized/",T), retimg = TRUE)
```

```{r}
image(as.nifti(roi.img), z = 21, plot.type='single')
```

```{r}
overlay.nifti(as.nifti(roi.img), as.nifti(mask.dat[,,1:110]), z = 21, plot.type = 'single')
```

### Median Filtering 

```{r}
filtered.img <- boss_filtering(roi.img, output_directory = paste0(getwd(), "/Filtered/",T), retimg = TRUE)
```

```{r}
overlay.nifti(as.nifti(filtered_img[,,1:110]), as.nifti(mask.dat[,,1:110]), z = 91, plot.type = 'single')
```
### Cell Tracking 

```{r}
setwd("/Users/eunchan/Box Sync/BOSS_example/Filtered/")


T1 <- readRDS("10-median-filtered-1.rds")
T2 <- readRDS("10-median-filtered-2.rds")
T3 <- readRDS("10-median-filtered-3.rds")
T4 <- readRDS("10-median-filtered-4.rds")
T5 <- readRDS("10-median-filtered-5.rds")
T6 <- readRDS("10-median-filtered-6.rds")
T7 <- readRDS("10-median-filtered-7.rds")
T8 <- readRDS("10-median-filtered-8.rds")
T9 <- readRDS("10-median-filtered-9.rds")
T10 <- readRDS("10-median-filtered-10.rds")
T11 <- readRDS("10-median-filtered-11.rds")
T12 <- readRDS("10-median-filtered-12.rds")


filtered.img <- array(0, dim=c(dim(T1),12))

filtered.img[,,,1] <- T1
filtered.img[,,,2] <- T2
filtered.img[,,,3] <- T3
filtered.img[,,,4] <- T4
filtered.img[,,,5] <- T5
filtered.img[,,,6] <- T6
filtered.img[,,,7] <- T7
filtered.img[,,,8] <- T8
filtered.img[,,,9] <- T9
filtered.img[,,,10] <- T10
filtered.img[,,,11] <- T11
filtered.img[,,,12] <- T12


```
```{r}
boss_tracking(img = filtered.img, output_directory = "~/Box Sync/BOSS_example/BOSS_result", maxZ = 110, maxT = 12)
```

### Visualization

```{r}
```

```{r setup}
library(bossR)
```

```{r}

```
